### 2013-06-23 outline.tex
Move outline out of manuscript

\documentclass{article}
\usepackage{bib/aas_macros}
\input{ligo-acronyms/acronyms}

\begin{document}

\section*{Outline}

\begin{enumerate}
\item Introduction {
	\begin{enumerate}
	\item Background: \acp{GRB}, \acp{CBC}, \acs{LIGO}
	\item Time scales of \acp{GRB}, \acp{CBC}
	\item Current sky localization algorithms and response times
	\end{enumerate}
}
\item Preliminaries {
	\begin{enumerate}
	\item Fisher information, \ac{CRLB}
	\item Matched\nobreakdashes-filter estimator
	\end{enumerate}
}
\item Algorithm {
	\begin{enumerate}
	\item Assumptions
	\item Signal model
	\item Likelihood function
	\item Priors and marginalization
	\item Procedure
	\end{enumerate}
}
\item Results {
	\begin{enumerate}
	\item Injection population
	\item Cumulative fraction vs. confidence level plots
	\item Sky localization accuracy as a function of distance or \ac{SNR}
	\item Speed
	\end{enumerate}
}
\item Discussion {
	\begin{enumerate}
	\item Potential further refinements
	\item What astrophysical targets were accessible before
	\item What new targets are available with the response time of seconds
	\item Lay out timeline of follow\nobreakdashes-up campaign, including different tiers of inference and different types of telescopes
	\end{enumerate}
}
\item Appendices {
	\begin{enumerate}
	\item Code listing?
	\item Any proofs that are needed in the body but would interrupt the text?
	\end{enumerate}
}
\end{enumerate}

\end{document}